---
title: "FluAnalysis"
author: "Aidan Troha"
editor: visual
---
We begin by using the library function to be able to use the `tidyverse` packages
```{r}
library(tidyverse)
```
We use the `here()` function from the `here` package to identify the file location and use `readRDS` to import the `.Rda` file.
```{r}
fludat_raw <- here::here("fluanalysis","data","raw_data","SympAct_Any_Pos.Rda")
flu_raw <- readRDS(fludat_raw)
```
We use the `anyNA()` function to indicate whether there are any missing values in the original data set. Also, we can use the `!c()` to make a subset of all the variables we do not want to include. On top of this, we can use `dplyr`'s `select()` to choose the variables we want to include in the new data set. Finally, we apply `drop_na()` to exclude any missing data. At the end, we validate that out analyses excluded all missing data using `anyNA()` on the new data set.
```{r}
anyNA(flu_raw)
flu_clean <- flu_raw %>%
                select(!c(contains(c("Score","Total","FluA","FluB",
                                     "Dxname","Activity")),"Unique.Visit")) %>%
                drop_na()
anyNA(flu_clean)
```
Again, we use the `here()` function to clearly indicate where the `RDS` file should be saved. We use `saveRDS()` to save the data set as the proper file type.
```{r}
fludat_clean <- here::here("fluanalysis","data","processed_data","flu_processed")
saveRDS(flu_clean,file=fludat_clean)
```
